---
title: "data.table"
date: 2017-03-02T15:45:27+01:00
draft: false
categories: ["R"]
tags: ["R", "data.table"]
---

## 1. What is data.table and why would you use it? 

* data.table is an R packge which let's you work on tabular datasets quickly and easily;

* comparing to base R or [dplyr](http://tomis9.com/tidyverse/#/dplyr) it's significantly faster;

* data.table has a concise and SQL-like syntax.

## 2. Basic functionalities 

### Creating a data.table 

```{r, message = FALSE}
library(data.table)

df <- data.frame(x = c("b","b","b","a","a"),
                 v = rnorm(5))

dt <- data.table(x = c("b","b","b","a","a"),
                 v = rnorm(5))
```

is exactly the same as creating a data.frame. The method `as.data.table()` works exaclty the same as `as.data.frame()`.

### Filtering 

Let's create a sample dataset first, baased on mtcars table: 
```{r}
sample_dataset <- as.data.table(datasets::mtcars)
```
Yes, you already have *datasets* package installed.

```{r}
sample_dataset[cyl == 6]
```

What happened? We chose only those cars, which have 6 cylinders. Data.table already knew that we mean a column named `cyl`, not an object from outside of the square brackets.

### Selecting columns 

```{r}
sample_dataset[, .(mpg, cyl, disp)][1:5]
```

What happened here?

* we used a special fucntion from data.table package: `.()`, which works just like vectors, but inside data.tables square brackets it treats columns as separate objects, so to work on column `mpg`, you simply type `mpg` instead of `"mpg"` or `sample_dataset$mpg`

* in square brackets we first provided a comma, as the first argument is always filtering. If we want to skip filtering, we simply write a comma;

* we chose the first five elements from our dataset. We could write even more square brackets after the whole statement and it would work as a pipe, but this would be too dplyr-ish.

### Grouping 

```{r}
sample_dataset[, .(mean_mpg = mean(mpg), count = .N), cyl]
```
* group by is the last statement inside the square brackets. In the example above, we group by column cyl;

* in the select clause we do exactly the same thing as in SQL statements;

* `.N` means *number of* or simply *count*.

### Reading and writing data 

data.table has the fastest reading and writing functions available in R. These are:

```{r}
fwrite(x = mtcars, file = 'mtcars.csv')
ds <- fread(file = 'mtcars.csv')
```

`fread` is pretty clever. It recognises if a file has headers, columns datatypes and separators. What I like the most in these functions is that I literally *never* have to provide any details about the file. Object and file names are always enough for data.table.

### Ordering data 

Very easy.

```{r}
sample_dataset[order(-gear, cyl)][1:5]
```

### Updating data 

In order to update our dataset we use the `:=` operator:

```{r}
sample_dataset[mpg > 30, carb := -1]
```

### Creating a new column 

In the same way as updating we can create a new column in place:

```{r}
sample_dataset[, new_column := 0]
print(sample_dataset[1:5])
```

But we don't have to do it in place:

```{r}
sample_dataset[, .(mpg, cyl, new_column2 = 0)][1:5]
```


## 3. Subjects still to cover: 

* `.I`(TODO)

* `.SD` + lapply (TODO)

* `merge()` (TODO)

* `setkey()` (TODO)
