---
title: "model validation"
date: 2019-05-23T12:46:03+02:00
draft: false
categories: ["Machine learning"]
tags: []
---

## 1. What is model validation and why would you do it?

You learned your model and naturally you are wondering how good it is. There are several ways to find out, like measuring the accuracy of predictions, but you may also want to check where exactly particular predictions come from, as this is far from obvious for "black box" models.

## 2. Examples

### cross-validation 

### confusion matrix 

https://www.rdocumentation.org/packages/caret/versions/3.45/topics/confusionMatrix

Confusion matrix is confusing at all as the name may suggest. 

* base R

- dataset preparation (described in more detail [here](http://tomis9.com/useful_processing))
```{r}
# prepare the dataset
library(caret)
species <- c("setosa", "versicolor")
d <- iris[iris$Species %in% species,]
d$Species <- factor(d$Species, levels = species)
trainIndex <- caret::createDataPartition(d$Species, p=0.7, list = FALSE, 
                                         times = 1)
train <- d[trainIndex,]
test <- d[-trainIndex,]
y_test <- test$Species == species[2]
```

and the logistic regression itself:
```{r}
m <- glm(Species ~ Sepal.Length, train, family = "binomial")
y_hat_test <- predict(m, test[,1:4], type = "response") > 0.5
```

We've prepared our predictions, as well as testing target, as vectors of binary values:
```{r}
y_test[1:10]
y_hat_test[1:10]
```

so now we may use a simple `table()` function to create a confusion matrix:
```{r}
table(y_hat_test, y_test)
```

* R caret

`caret` provides a much broader summary of confusion matrix:
```{r}
library(caret)
m2 <- train(Species ~ Sepal.Length, train, method = "glm", family = binomial)
confusionMatrix(predict(m2, test), test$Species)
```

```{python, engine.path = '/usr/bin/python3'}
from sklearn.datasets import load_iris
from sklearn.model_selection import train_test_split
from sklearn.linear_model import LogisticRegression
from sklearn.metrics import accuracy_score, confusion_matrix

iris = load_iris()

cond = iris.target != 0
X = iris.data[cond]
y = iris.target[cond]

X_train, X_test, y_train, y_test = train_test_split(
    X, y, test_size=0.33, random_state=42)

lr = LogisticRegression()

lr.fit(X_train, y_train)
accuracy_score(lr.predict(X_test), y_test)
print(confusion_matrix(lr.predict(X_test), y_test))
```


### ROC, AUC 

Let's say you created a simple classifier, e.g. using logistic regression. The classifier does not return classes though, but the probability that this particular observation belongs to class 1. As what we need are classes, not probabilities, we have to somehow *map* these probabilities into classes. The easiest way to achieve this is by using a function like:

$$ f(t) = \begin{cases} 1 & \text{when $p \geqslant t$} \\ 0 & \text{when $p < t$} \\ \end{cases} $$

where $t$ is a threshold set by *you*. Choose wisely ;)

Choosing a proper value of $t$ is known as a "Precision / Recall tradeoff":

- [a wonderful, intuitive article](https://towardsdatascience.com/precision-vs-recall-386cf9f89488)

- [another good article](https://towardsdatascience.com/beyond-accuracy-precision-and-recall-3da06bea9f6c)

- [A wonderful article about AUC and ROC curves](https://towardsdatascience.com/understanding-auc-roc-curve-68b2303cc9c5). There is no nedd to duplicate it.

Different values of TPR and FPR for various $t$ create a ROC curve. Area Under this Curve is called AUC.

*R - using ROCR package*
```{r, message = FALSE}
library(ROCR)
plot_roc_get_auc <- function(pred, test_labels) {
  roc_pred <- ROCR::prediction(pred, test_labels)
  roc_perf <- ROCR::performance(roc_pred, measure = "tpr", x.measure = "fpr")
  ROCR::plot(roc_perf, col = 1:10)
  abline(a = 0, b = 1)
  auc_perf <- ROCR::performance(roc_pred, measure = "auc", x.measure = "fpr")
  return(auc_perf@y.values[[1]])
}

species <- c("setosa", "versicolor")
iris_bin <- iris[iris$Species %in% species,]
iris_bin$Species <- factor(iris_bin$Species, levels = species)
trainIndex <- caret::createDataPartition(iris_bin$Species, p=0.7, list = FALSE, 
                                         times = 1)
train <- iris_bin[trainIndex,]
test <- iris_bin[-trainIndex,]


m <- glm(Species ~ Sepal.Length, train, family = binomial)
plot_roc_get_auc(
  pred = predict(m, test[,1:4], type = "response"),
  test_labels = as.integer(test[["Species"]]) - 1)


rf <- randomForest::randomForest(Species ~ ., data = train)
plot_roc_get_auc(
  pred = predict(rf, test[, 1:4], type = "prob")[,'versicolor'],
  test_labels = as.integer(test[["Species"]]) - 1)

```


https://towardsdatascience.com/supervised-machine-learning-model-validation-a-step-by-step-approach-771109ae0253

ROC - TPR vs FPR, where

$$ \textrm{TPR} = \frac{\textrm{TP}}{\textrm{TP} + \textrm{FP}} $$

TPR - True Positive Rate
TP - True Positive
FP - False Positive


parameter search

https://scikit-learn.org/stable/modules/grid_search.html#tuning-the-hyper-parameters-of-an-estimator

TODO: https://rviews.rstudio.com/2019/03/01/some-r-packages-for-roc-curves/

TODO: https://www.saedsayad.com/model_evaluation_c.htm

TODO: https://eli5.readthedocs.io/en/latest/

TODO: https://towardsdatascience.com/explaining-feature-importance-by-example-of-a-random-forest-d9166011959e

TODO: https://github.com/slundberg/shap
